```{r include = FALSE, message = FALSE}
source("../source/analysis.R")
```


# ---------- General Overview of `incarceration_trends` and `jail_jurisdiction` ----------

For this assignment, I will be comparing data drawn from `incarceration_trends` and  
`jail_jurisdiction`, observing any differences between demographics such as race and
gender. The information recorded in `incarceration_trends` is based on incarcerated
populations & the surrounding areas on a county-level basis. `jail_jurisdiction` is
based on incarcerated populations in local county jails.

Note:
Jail refers to individuals who are held in lawful custody (i.e. have been convicted
of a minor offense or are waiting to go on trial for their accusation) for a short
period of time. The word jail implies that it is a smaller, local facility. Prison
refers to individuals who have been convicted of more serious offenses and are
sentenced to longer terms. Prison can also be referred to as "state prison" or
"federal prison", and refers to a larger facility of confinement than a county jail.


This table summarizes general information about each county:
```{r echo = FALSE, message = FALSE}
head(county_info)
```


# ---------- Annual Total Incarcerations ----------

This section contains charts that show incarceration statistics for gender and race
in jail and prison annually. The first columns include the numbers of the general
population, which can be compared to the number of incarcerated individuals and
provided demographics of that population.

Charts that compare gender (either for jail or prison) include total numbers of each
gender and age categories. `jail_sortgender` contains more data than its
`prison_sortgender` counterpart, as more data was provided by the Vera Institute. The
data is divided up not only by gender but also by "juvenile" or "adulthood".


`jail_sortgender`
```{r echo = FALSE, message = FALSE}
head(jail_sortgender)
```


`prison_sortgender`
```{r echo = FALSE, message = FALSE}
head(prison_sortgender)
```




Similarly, `jail_sortrace` and `prison_sortrace` compare numbers categorized by race.
The total incarcerated population is grouped by race, providing a total population
and individuals who are between the ages of 15 to 64. `prison_sortrace` was able to
provide further information by specifying the gender of individuals along with race.


`jail_sortrace`
```{r echo = FALSE, message = FALSE}
head(jail_sortrace)
```


`prison_sortrace`
```{r echo = FALSE, message = FALSE}
head(prison_sortrace)
```


# ---------- Comparing Jail Incarceration Statistics ----------

`jail_genderstats` summarizes information from `incarceration_trends` which compares
new admissions to jail and the population percent increase, per their respected gender.
`jail_racestats` similarly compares incarcerated individuals by race, and takes data
from the earliest recorded year and the latest recorded year. Numbers include the
percent change (either increase or decrease) in population numbers for racial groups.
Also includes the number of individuals discharged that year.


`jail_genderstats`
```{r echo = FALSE, message = FALSE}
head(jail_genderstats)
```


`jail_racestats`
```{r echo = FALSE, message = FALSE}
head(jail_racestats)
```


# ---------- Comparing Prison Incarceration Statistics ----------

`prison_genderadm` compares newly admitted inmates in state/federal prisons by gender
over time. Grouped by year, the chart shows the sum of every new inmate in the country
and the percent increase along with it. `prison_raceadm` compares newly admitted inmates
in state/federal prisons by race over time. The chart takes into account total
incarcerated numbers for each race, along with general admission numbers & rates and
female/male admits & admission rates.


`prison_genderadm`
```{r echo = FALSE, message = FALSE}
head(prison_genderadm)
```


`prison_raceadm`
```{r echo = FALSE, message = FALSE}
head(prison_raceadm)
```


# ---------- Trends Over Time Chart ----------

This graph compares the sum of inmates in the country by race over time. I chose to
compare this because I thought it would be important to see the racial distribution of
the inmate population. A lot of factors play into these numbers, including minorities
who suffer from racism and false accusations, so it's interesting to see how that
impacts the incarcerated populations compared to white individuals. Even though black
individuals make up a lower percentage of racial demographics in the U.S. compared to
their white counterparts, their incarcerated population is still similarly as high.

```{r echo = FALSE, message = FALSE}
ggplot(
  melt_totc, 
  aes(
    x = Year, 
    y = `New Admits`, 
    group = Race, 
    color = Race
  )) +
  geom_line() +
  ggtitle("New Admits by Race Over Time")
```


# ---------- Variable Comparison Chart ----------

This graph compares the racial distribution in jail for the most recent year of data.
Disclaimer: there are no recordings of racial groups in the years 2017 and 2018 for
Latinx, AAPI, Native American, and other. Despite that, the numbers have a similar
pattern to the trends over time chart; black individuals are still comparatively high
than other minorities, despite the overall racial distribution of races in the U.S.
Second to white individuals, which made up ~47% of the incarcerated population, black
individuals made up ~34% while other minorities are much lower (referencing to other
years where the percentages of other minority groups were made available).

```{r echo = FALSE, message = FALSE}
ggplot(
  melt_jailmax,
  aes(x = Race, y = Percent)) +
  geom_bar(stat = "identity") +
  ggtitle(paste("Racial Distribution of Incarcerated Population (", melt_jailmax[1, "Year"], ")", sep = "")
)
```

# ---------- Map ----------

This map shows how races are distributed geographically from 2010. I picked this year
instead of the most recent year of recorded data because the numbers between racial
groups existed and were more comparable (frame of reference was the trends over time
chart). I divided each map into racial groups to make comparing between them easy.
Although numbers for AAPI and Native Americans are higher, they're still relatively
low compared to white, black, and latinx individuals. From the map, it appears that
those incarcerations are more prevalent towards the south, especially in states like
Texas and Florida. Higher numbers also showed up in California. 

```{r echo = FALSE, message = FALSE}
ggplot(state_jail_join) +
  geom_polygon(
    mapping = aes(
      x = long, 
      y = lat, 
      group = group, 
      fill = `Population`
    )
  ) +
  coord_map() +
  scale_fill_continuous(high = "#fff44f", low = "#034746") +
  facet_grid(Race ~ .) + # creates maps for each race
  labs(fill = "Population") +
  ggtitle(paste("Racial Distribution in Jail (2010)"))
```